Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2016-05-13 ms.tex
another round of casey comments

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observables of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ the new VIRUS integral field unit (IFU) spectrograph, over 420 \degsq\ on the sky with a 1/4.5 fill factor. The instrument covers the blue/optical portion of the spectrum (3500--5500~\AAA), allowing surveys with VIRUS to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) IFU surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. Using both observing strategies, we produce cluster mass probability density functions $P(M_{200c}|\vec{x})$ of a cluster mass ($M_{200c}$) as a function of an observable parameter or set of observable parameters, $\vec{x}$, where we consider cases of where $\vec{x} = \{\sigma\}$, $=\{\sigma, z\}$, and $=\{\sigma,z, N_{gal}\}$ for velocity dispersion, $\sigma$, redshift, and number of observed galaxies, $N_{gal}$. \editorial{This sentence will summarize how well HETDEX will do, and how that might be improved with Targeted observations.}
\end{abstract}

@@

The second broad camp is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses direct observations of a small set of clusters and then uses known mass estimators, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, cosmic microwave background (CMB) observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable--mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of CMB photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters, and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

\subsection{The Buzzard Mock Catalogs}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. The [\ion{O}{ii}] luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.} \label{fig: oii sdss} 
\end{figure*}

@@

\subsection{Conditional [\ion{O}{ii}] Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{ii}] emission line strengths to the galaxies in the Buzzard catalog. We use 503113 galaxies from $z = 0.05 - 0.2$, which are selected with no redshift warning, and place each galaxy on a color-magnitude diagram (CMD) of $M_r$ and $g-r$; see the left panel of Figure \ref{fig: oii sdss}.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $p(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{\url{http://astlib.sourceforge.net}}}:
\begin{equation}
C_{BI} = M + \frac{\sum_{|u_i| < 1} (x_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $x_i$ are the individual redshifts, $M$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(x_i - M)}{c\mathrm{MAD}}
\end{equation}.
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $c$ is the a tuning constant. We choose $c=6$ (the suggested value) which balances computational speed and location accuracy.

@@

We follow the maximum likelihood method of \cite{Walk2006} to estimate the line-of-sight velocity dispersion (LOSVD). We maximize the probability function 
\begin{equation}\label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_p^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_p^2)}\biggr]
\end{equation}
where $\sigma_p$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs respectively, using a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_p < 1400$ \kms. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster. When the full distribution of LOSVDs is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the 16th and 84th percentiles of the same distribution.

@@

\subsubsection{Other Estimates of Dynamical Mass -- Introduction}
In this section we use two methods to predict the mass of a cluster based on other observables. Often the cluster mass is estimated based on a single observable, X-ray temperature, LOSVD, richness and others (see Section~\ref{sec: Introduction} for referenced examples). Here we combine many observables to attempt to correct the mass inferred solely from the velocity dispersion. The first method is traditional ``probability based'' where we marginalize over a series of observables to find the most probable cluster mass. The second is based on a machine learning (ML) algorithm which attempts to infer the relationship between the observables and the desired output, the cluster mass. Both of these methods are examples of supervised learning algorithms where the relationship between the observable parameters and the target parameter (the cluster mass) are both known.

@@

\subsubsection{Probability Based}\label{sec:probability method}
For internal consistency between this and the ML based method we use 70\% of the clusters to establish a conditional probability density of $p(M_{200c} | \vec{x})$ which we then use as the mass probability density for the remaining clusters with similar features. In this way we, ``train'' the probability density using the existing simulated data, and apply it to the ``test'' sample (the remaining 30\% of the data not used as the ``training'' set).

Figure~\ref{fig: probability corner} shows all of the one (marginalized probability) and two (joint probability) dimensional projections of the posterior probability distributions of the features of the training data.

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/prob_example.pdf} 
	\caption{Corner plot of the \emph{training} data with features $\sigma$ and $z$. The corner plots shows all of the one and two dimensional posterior probability distributions used to determine the correct cluster mass. The colored rectangles, which have been enlarged for clarity, show the slices needed to create a conditional probability distribution of the mass, $P(M_{200c}|\vec{x})$. See text for a complete description. } \label{fig: probability corner} 
\end{figure*}

The conditional probability of the mass $P(M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ is determined by taking a slice through the joint probability distributions in bins centered on the desired value. The slices, shown by the colored bars in Figure~\ref{fig: probability corner}, are centered on $\sigma = 400$ \kms and $z=0.17$. The distribution of cluster mass contained in the three dimensional bin given by the intersection of these slices is $P(M_{200c}|\vec{x} = \{\sigma=400 \kms,z=0.17\})$.

\begin{figure*}
	\includegraphics[width=0.8\textwidth]{figures/recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and true cluster mass for the Targeted and Survey observing strategies. We have applied HETDEX-like observational limits on the cluster galaxy detection, and require at least five galaxies to be detected for a cluster to be recovered. The solid lines are the median values and the shaded regions represent the 68\% scatter. The significant decline in galaxies observed with the Survey strategy is due to gaps in the VIRUS IFU.} \label{fig: recovery} 
\end{figure*}

@@

The prediction intervals here are based on the general method of quantile regression forests \citep{Meinshausen2006}. The general idea is that all response variables are recorded, not just the mean. Then the prediction can be returned as the full conditional probability distribution of all responses, which allows us to generate the prediction intervals. The 68\% prediction interval is determined by calculating the 16th and 84th percentile of the full conditional probability distribution.

@@

\subsection{Recovery of Cluster Members}
As discussed in Section \ref{sec: observations} the observational constraints place limits on the total number of clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or Targeted observations. Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. We require a detection of ($N_{obs} >=5$) galaxies for a cluster to be “detected” and included in this figure. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. With Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses.

For the clusters recovered as a function of redshift (left panel of Figure \ref{fig: recovery}), there are two effects at work. The decrease in recovery fraction with increasing redshift is a magnitude effect. We check this by limiting the cluster galaxy detection by absolute magnitude which increases the recovery fraction to $>70\%$ at all redshifts, implying the decline is a result of the apparent magnitude cluster galaxy detection threshold. The second key feature is the strong decline in clusters recovered from Survey observations. This is due to gaps in the VIRUS IFU. The median recovery fraction in Survey observations is almost exactly 4.5 times less than the Targeted median recovery fraction. As the total filling factor of the Survey increases the two lines will converge.

The recovery rate as a function of cluster mass, right panel of Figure \ref{fig: recovery}, shows that of the the low mass clusters we detect ($N_{obs} >5$), we observe the majority of the galaxies. This also shows a rapid decrease in the detection fraction, which can again be explained by considering absolute magnitudes instead of apparent magnitudes, as above. Also, high mass clusters are rare, so in order to detect them we must probe a large volume of space. The higher redshift cluster members suffer from the limiting apparent magnitude and suppress the recovery fraction at fixed mass. If we were to limit the Survey to $z<0.2$ we find the recovery fraction of clusters, at all masses, increases substantially, and we find a much more consistent detection fraction across all masses.

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The solid, colored line is the median predicted mass for the Targeted observing, and the colored, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.} \label{fig:Probability comparison} 
\end{figure*}

In this section we discuss the how accurately we recover the true cluster mass from a set of observations. We report on three methods, the power law based approach (Eq. \ref{eq:power law}), the probability based approach (Section \ref{sec:probability method}) and the ML based method (Section \ref{sec:machine learning method}). For each method we consider observations with Perfect knowledge, Targeted observations and Survey observations.

Because it represents the best possible scenario, the Perfect knowledge observations should serve as baseline to compare the power law based, probability based and ML cluster mass recovery methods. And, while there are many possible metrics to evaluate performance, we compute two: the average bias (given in Table \ref{tbl:mass bias})
\begin{equation}
\mathrm{\mu_{bias}}(y,\hat{y}) = \frac{1}{N} \sum_{i=1}^N (\hat{y_i} - y_i).
\end{equation}
where $y$ are the true values and $\hat{y}$ are the predicted values, and the scatter about the bias (given in Table \ref{tbl:mass scatter})
\begin{equation}
	\sigma_{bias}(y,\hat{y}, \mu_{bias}) = \bigg{[}\frac{1}{N-1} \sum_{i=1}^N (\hat{y_i} - y_i - \mu_{bias})^2 \bigg{]}^{1/2}
\end{equation}
with $N$ clusters in a given bin. Both metrics evaluate how closely the ensemble of predicted cluster masses are to the true cluster masses, and in both cases values closer to zero are better.

We begin with the Perfect knowledge observations. These observations are of the same clusters as the Targeted observations but without any observational limits. The cluster masses predicted by Equation \ref{eq:power law} gives the following results. For clusters with masses between Log M/\Msol $= 13 - 15.5$, we find $\mu_{bias} = 0.15\pm{0.002}$ dex and $\sigma_{bias} = 0.24\pm{0.001}$. The scatter in recovered masses can be attributed to both physical and numerical effects. The presence of any in-falling matter onto lower mass clusters can introduce a significant amount of substructure, leading to artificial biasing of the measured LOSVD to higher values, increasing the predicted mass \citeeg{Ntampaka2015}. Also, as the number of cluster galaxies decreases the LOSVD PDF is poorly sampled leading to poorly recovered cluster masses due to numerical effects. The cluster masses presented here are recovered using the best possible conditions, where we have Perfect knowledge of the cluster membership. In reality, the mass recovery levels presented in this section represent an upper bound (the best) on the accuracy achievable through this method.

@@

For the Targeted and Survey observations the power law predicted cluster masses give $\mu_{bias} =0.13\pm{0.003}$ dex, $\sigma_{bias} = 0.37\pm{0.002}$ and $\mu_{bias} =0.11\pm{0.009}$ dex, $\sigma_{bias} = 0.36\pm{0.006}$ , respectively. So for the clusters that we detect with Survey observations, we obtain similar levels of accuracy as to the Targeted observations, on the average. This does not mean that the Survey observations cannot be improved by Targeted observations. In fact, when comparing only the galaxies which overlap between the two samples the bias and scatter of the Targeted observations is significantly decreased as more cluster member galaxies are detected, better sampling the LOSVD PDF. The Targeted observations performs similarly on the average because many lower mass clusters are included in the sample, increasing the bias of the overall sample.

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the ML based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The solid, colored line is the median predicted mass for the Targeted observing, and the colored, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.} \label{fig: ML comparison} 
\end{figure*}

>>>>

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observables of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion using probability based and machine learning methods. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ new Visible Integral-Field Replicable Unit Spectrographs (VIRUS), over 420 \degsq\ on the sky with a 1/4.5 fill factor. The instrument covers the blue/optical portion of the spectrum ($3500-5500~\AAA$), allowing surveys with VIRUS to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) IFU surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. Using both observing strategies, we produce cluster mass probability density functions $P(M_{200c}|\vec{x})$ of a cluster mass ($M_{200c}$) as a function of an observable parameter or set of observable parameters, $\vec{x}$, where we consider cases of where $\vec{x} = \{\sigma\}$, $=\{\sigma, z\}$, and $=\{\sigma,z, N_{gal}\}$ for velocity dispersion, $\sigma$, redshift, and number of observed galaxies, $N_{gal}$. \editorial{This sentence will summarize how well HETDEX will do, and how that might be improved with Targeted observations.}
\end{abstract}

@@

The second broad camp is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses direct observations of a small set of clusters and then uses known mass estimators, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable--mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

\subsection{The Buzzard Mock Catalogs}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503,113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. The [\ion{O}{ii}] luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.} \label{fig: oii sdss} 
\end{figure*}

@@

\subsection{Conditional [\ion{O}{ii}] Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{ii}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 galaxies from $z = 0.05 - 0.2$, which are selected with no redshift warning, and place each galaxy on a color-magnitude diagram (CMD) of $M_r$ and $g-r$; see the left panel of Figure \ref{fig: oii sdss}.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $p(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{\url{http://astlib.sourceforge.net}}}:
\begin{equation}
C_{BI} = M + \frac{\sum_{|u_i| < 1} (x_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $x_i$ are the individual redshifts, $M$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(x_i - M)}{c\mathrm{MAD}}.
\end{equation}
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $c$ is the a tuning constant. We choose $c=6$ (the suggested value) which balances computational speed and location accuracy.

@@

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). We maximize the probability function 
\begin{equation}\label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
where $\sigma_{1D}$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs (which we have assumed to be normally distributed) respectively, using a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster. When the full distribution of LOSVDs is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the 16th and 84th percentiles of the same distribution.

@@

\subsubsection{Other Estimates of Dynamical Mass -- Introduction}
In the following sections we use two methods to predict the mass of a cluster based on other observables. Often the cluster mass is estimated based on a single observable, X-ray temperature, LOSVD, richness and others (see Section~\ref{sec: Introduction} for referenced examples). Here we combine many observables to attempt to correct the mass inferred solely from the velocity dispersion. The first method is traditional ``probability based'' where we marginalize over a series of observables to find the most probable cluster mass. The second is based on a machine learning (ML) algorithm which attempts to infer the relationship between the observables and the desired output, the cluster mass. Both of these methods are examples of supervised learning algorithms where the relationship between the observable parameters and the target parameter (the cluster mass) are both known.

@@

\subsubsection{Probability Based}\label{sec:probability method}
For internal consistency between this and the ML based method we use 70\% of the clusters to establish a conditional probability density of $p(M_{200c} | \vec{x})$ which we then use as the cluster mass probability density for the remaining clusters with similar features. In this way we, ``train'' the probability density using the existing simulated data, and apply it to the ``test'' sample (the remaining 30\% of the data not used as the ``training'' set).

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/prob_example.pdf} 
	\caption{Illustration of the probability based cluster mass prediction method. \emph{Left}: The two dimensional posterior probability distribution of LOSVD and redshift used to determine the correct cluster mass. The pink and green rectangles show the locations of two example galaxies used to create the conditional probability distribution of the mass, $P(M_{200c}|\vec{x})$. \emph{Right}: The conditional probability distribution of the cluster mass for the two example galaxies. See text for a complete description.} \label{fig: probability corner} 
\end{figure*}

In this method, the cluster masses are predicted much in the same way we assign [\ion{O}{ii}] flux to the Buzzard catalog galaxies. Figure~\ref{fig: probability corner} illustrates the process. The left panel shows the two dimensional (joint probability) projections of the posterior probability distributions of the feature training data. The conditional probability of the cluster mass $P(M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ (shown in the right panel) is determined by selecting a region in the joint probability distribution. For example, using the LOSVD and redshift features we create $P(M_{200c}|\vec{x})$ for two test galaxies, shown by the green and pink boxes in the left panel of Figure~\ref{fig: probability corner}. These example galaxies have features $\vec{x} = \{\sigma=200\, \kms,z=0.16\}$ and $\vec{x} = \{\sigma=400\, \kms,z=0.36\}$. We select all galaxies in our training sample with similar features and create the conditional probability distributions shown in the right panel.

@@

The prediction intervals here are based on the general method of quantile regression forests \citep{Meinshausen2006}. The general idea is that all response variables are recorded, not just the mean. Then the prediction can be returned as the full conditional probability distribution of all responses, which allows us to generate the prediction intervals. We report the 68\% prediction interval as the second moment of the full conditional probability distribution.

@@

\subsection{Recovery of Cluster Members}
\begin{figure*}
	\includegraphics[width=0.8\textwidth]{figures/recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and true cluster mass for the Targeted and Survey observing strategies. We have applied HETDEX-like observational limits on the cluster galaxy detection, and require at least five galaxies to be detected for a cluster to be recovered. The shading indicates the mean recovery fraction for clusters within each small bin of redshift and cluster mass. We find a significant decrease in the recovery of galaxies with increasing redshift. This leads to lower recovery fractions of high mass clusters as many only exist at larger redshifts. The significant decline in the number of galaxies observed with the Survey strategy is due to gaps in the VIRUS IFU, where the galaxies are missed.} \label{fig: recovery} 
\end{figure*}

As discussed in Section \ref{sec: observations} the observational constraints place limits on the total number of clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or Targeted observations. Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. We require a detection of ($N_{obs} >=5$) galaxies for a cluster to be “detected” and included in this figure. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. With Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses.

For the clusters recovered as a function of redshift (y-direction of Figure~\ref{fig: recovery}), there are two effects at work. The decrease in recovery fraction with increasing redshift is a magnitude effect. Increasing redshift results in a lower galaxy recovery fraction. Because HETDEX is limited in apparent magnitude, we expect to recovery fewer galaxies at higher redshift. We confirm this by limiting the cluster galaxy detection by absolute magnitude, by observing all galaxies with $M_g < -11$. This increases the recovery fraction to $>70\%$ at all redshifts, implying the decline is a result of the apparent magnitude cluster galaxy detection threshold. The second key feature is the strong decline in clusters recovered from Survey observations. This is due to gaps in the VIRUS IFU. The median recovery fraction in Survey observations is almost exactly 4.5 times less than the Targeted median recovery fraction. As the total filling factor of the Survey increases the two lines will converge.

The recovery rate as a function of cluster mass, x-direction of Figure \ref{fig: recovery}, shows that of the the low mass clusters we detect ($N_{obs} >5$), we observe the majority of the galaxies. This also shows a rapid decrease in the detection fraction, which can again be explained by considering absolute magnitudes instead of apparent magnitudes, as above. Because there are few high mass ($M>5\times10^{14}$ \Msol) clusters, many of which are at moderately high redshift, the higher redshift cluster members suffer from the limiting apparent magnitude and suppress the recovery fraction at fixed mass. If we were to limit the Survey to $z<0.2$ we find the recovery fraction of clusters, at all masses, increases substantially, and we find a much more consistent detection fraction across all masses. Again perfect observations would detect all members, at all masses, so in this way, Figure~\ref{fig: recovery} is really just a statement about the magnitude limit and [\ion{O}{ii}] line-flux limit of the survey.

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation, in both panels, between $M_{200c}$ and $M_{pred}$. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.} \label{fig:Probability comparison} 
\end{figure*}

In this section we discuss the how accurately we recover the true cluster mass from a set of observations. We report on three methods, the power law based approach (Eq. \ref{eq:power law}), the probability based approach (Section \ref{sec:probability method}) and the ML based method (Section \ref{sec:machine learning method}). For each method we consider observations with Perfect knowledge, Targeted observations and Survey observations. The cluster masses presented here are recovered using the best possible conditions, where we have Perfect knowledge of the cluster membership. In reality, the mass recovery levels presented in this section represent an upper bound (the best) on the accuracy achievable through this method.

Because it represents the best possible scenario, the Perfect knowledge observations should serve as baseline to compare the power law based, probability based and ML cluster mass recovery methods. And, while there are many possible metrics to evaluate performance, we compute two: the average bias (given in Table \ref{tbl:mass bias})
\begin{equation}
\mathrm{\mu_{bias}}(y,y_{pred}) = \frac{1}{N} \sum_{i=1}^N (y_{pred,i} - y_i).
\end{equation}
where $y$ are the true values and $y_{pred}$ are the predicted values, and the scatter about the bias (given in Table \ref{tbl:mass scatter})
\begin{equation}
	\sigma_{bias}(y,y_{pred}, \mu_{bias}) = \bigg{[}\frac{1}{N-1} \sum_{i=1}^N (y_{pred,i} - y_i - \mu_{bias})^2 \bigg{]}^{1/2}
\end{equation}
with $N$ clusters in a given bin. Both metrics evaluate how closely the ensemble of predicted cluster masses are to the true cluster masses, and in both cases values closer to zero are better.

We begin with the Perfect knowledge observations. These observations are of the same clusters as the Targeted observations but without any observational limits. The cluster masses predicted by Equation \ref{eq:power law} gives the following results. For clusters with masses between Log M/\Msol $= 13 - 15.5$, we find $\mu_{bias} = 0.148\pm{0.008}$ dex and $\sigma_{bias} = 0.193\pm{0.001}$. The scatter in recovered masses can be attributed to both physical and numerical effects. The presence of any in-falling matter onto lower mass clusters can introduce a significant amount of substructure, leading to artificial biasing of the measured LOSVD to higher values, increasing the predicted mass \citeeg{Ntampaka2015}. Also, as the number of cluster galaxies decreases the LOSVD PDF is poorly sampled leading to poorly recovered cluster masses due to numerical effects.

@@

For the Targeted and Survey observations the power law predicted cluster masses give $\mu_{bias} =0.135\pm{0.003}$ dex, $\sigma_{bias} = 0.370\pm{0.002}$ and $\mu_{bias} =0.148\pm{0.008}$ dex, $\sigma_{bias} = 0.324\pm{0.006}$ , respectively. So for the clusters that we detect with Survey observations, we obtain similar levels of accuracy as to the Targeted observations, on the average. This does not mean that the Survey observations cannot be improved by Targeted observations. In fact, when comparing only the galaxies which overlap between the two samples the bias and scatter of the Targeted observations is significantly decreased as more cluster member galaxies are detected, better sampling the LOSVD PDF. The Targeted observations performs similarly on the average because many lower mass clusters are included in the sample, increasing the bias of the overall sample.

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the ML based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.} \label{fig: ML comparison} 
\end{figure*}
